Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2020-10-30 paper/main.tex
comments

Moreover, we consider real LOP datasets from the LOPLIB repository. \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
In particular, we consider the following types: IO, Random instances of type A and Random instances of type B.

@@

The plot shows that UMM is indeed very sensitive to the values of $r_1$ and
$r_2$, yet, low values of $r_1$ and high values of $r_2$ almost always produce
the best results, independently of the problem. Nevertheless, there are a few
cases where an interaction seems to occur and particular combinations perform
remarkably well.\MANUEL{Missing data, whitespaces}  In particular, we observe that the pattern on
the two left-most figures differs from the two right-most ones. While on the
left the fitness gets worse homogeneously from the top right corner to the
bottom left one, the pattern on the right is not as homogeneous, yet still
fitness gets worse with decreasing $r_2$.  We observe that this behavior is
related to the adequacy of UMM to the problem: the homogeneous behavior of
parameters $r_1$ and $r_2$ depends on the problem (not to the particular
instances) and the more homogeneous the behavior, the better performance of
UMM. We elaborate this argument in the following sections with detail.

@@

On the other hand, % \EKHINE{cambio el orden}
CEGO starts with a random sampling
of 10 solutions, with are quite poor as expected, followed by building and
updating the Kriging model based new evaluations, which leads to an impressive
improvement in less than 50 evaluations. However, progress is %more or less 
halted at this point and the model seems to have trouble generating better
solutions. Lower uniformity values (e.g., $\phi=.5$) show worsening evaluations
despite the increase in solutions used to build the model. In summary, the
Kriging model clearly helps CEGO to quickly identify good solutions in very few
evaluations. However, for some reason, CEGO is not able to further improve
those solutions given more data. This may due to the model not being able to
produce accurate predictions or the underlying optimizer not being able to find
improved solutions using the model.\MANUEL{perhaps we should measure the
  prediction error and see if it grows or decreases. That would answer this
  question.}

@@

In the following paragraphs we analyze the result over real rata. Fig.~\ref{fig:loplib} shows the results on the instances available in LOPLIB.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}} The distribution of this instances differs from the symthetic LOP instances, and thus so does the landscape. However, UMM shows a very similar performance: there is a (close to) linear decrease on the function evaluations of new solutions along iterations. This suggests that UMM very good fits the fitness landscape. Regarding CEGO, we can take similar conclussions for the first 50-100 iterations. However, the processes had to be halted after hitting the wall time, so no conlusion can be taken for further iterations. Overall, the improvement from the first (random) to the last solutions is 2-3 times larger in UMM than CEGO.

In the case of the PFSP, none of the algorithms are able to achieve a progress
better than a random search would, as illustrated by the results
(Fig.~\ref{fig:rec05}) obtained on \texttt{rec05}, which is the smallest PFSP
instance considered here. The lack of any apparent convergence suggests that
the models (both Bayesian and probabilistic) are not learning anything about
the fitness landscape. Results for larger PFSP corroborate this conclusions
(\supplement). Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
that other distance measures are more suited than Kendall's-$\tau$ distance for
the PFSP, the reported fitness differences between various distance metrics are
small, hence, we conjecture that the same behaviour will be observed with other
distance metrics.

@@

Table~\ref{tab:results} shows overall results over all instances evaluated (excluding the synthetic LOP instances). In particular, the mean fitness (and standard deviation) of the best solution found at the end of each run,
In terms of best fitness achieved on each run, and taking into account the
random behaviour observed in some instances and discussed above,\ekhine{}{falta un cacho aqui}

>>>>

Moreover, we consider real LOP datasets from the LOPLIB repository.\footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
In particular, we consider the following types: IO, Random instances of type A and Random instances of type B.

@@

The plot shows that UMM is indeed very sensitive to the values of $r_1$ and
$r_2$, yet, low values of $r_1$ and high values of $r_2$ almost always produce
the best results, independently of the problem. Nevertheless, there are a few
cases where an interaction seems to occur and particular combinations perform
remarkably well.\MANUEL{Missing data, whitespaces where more than 6 hours of wall-clock time}  In particular, we observe that the pattern on
the two left-most figures differs from the two right-most ones. While on the
left the fitness gets worse homogeneously from the top right corner to the
bottom left one, the pattern on the right is not as homogeneous, yet still
fitness gets worse with decreasing $r_2$.  We observe that this behavior is
related to the adequacy of UMM to the problem: the homogeneous behavior of
parameters $r_1$ and $r_2$ depends on the problem (not to the particular
instances) and the more homogeneous the behavior, the better performance of
UMM. We elaborate this argument in the following sections with detail.\MANUEL{top and bottom}

@@

On the other hand, % \EKHINE{cambio el orden}
CEGO starts with a random sampling
of 10 solutions, with are quite poor as expected, followed by building and
updating the Kriging model based new evaluations, which leads to an impressive
improvement in less than 50 evaluations. However, progress is %more or less 
halted at this point and the model seems to have trouble generating better
solutions. Lower uniformity values (e.g., $\phi=.5$) show worsening evaluations
despite the increase in solutions used to build the model. In summary, the
Kriging model clearly helps CEGO to quickly identify good solutions in very few
evaluations. However, for some reason, CEGO is not able to further improve
those solutions given more data. This may due to the model not being able to
produce accurate predictions or the underlying optimizer not being able to find
improved solutions using the model.
%\MANUEL{perhaps we should measure the
%  prediction error and see if it grows or decreases. That would answer this
%  question.}

@@

  In the following paragraphs we analyze the result over real rata. Fig.~\ref{fig:loplib} shows the results on the instances available in LOPLIB.
  %% MANUEL: Ya dimos la URL arriba, así que no la repetiría aquí
  % \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
  The distribution of this instances differs from the symthetic LOP instances, and thus so does the landscape. However, UMM shows a very similar performance: there is a (close to) linear decrease on the function evaluations of new solutions along iterations. This suggests that UMM very good fits the fitness landscape. Regarding CEGO, we can take similar conclussions for the first 50-100 iterations. However, the processes had to be halted after hitting the wall time, so no conlusion can be taken for further iterations. Overall, the improvement from the first (random) to the last solutions is 2-3 times larger in UMM than CEGO.

In the case of the PFSP, none of the algorithms are able to achieve a progress
better than a random search would, as illustrated by the results
(Fig.~\ref{fig:rec05}) obtained on \texttt{rec05}, which is the smallest PFSP
instance considered here. The lack of any apparent convergence suggests that
the models (both Bayesian and probabilistic) are not learning anything about
the fitness landscape. Results for larger PFSP corroborate these conclusions
(\supplement). Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
that other distance measures are more suited than Kendall's-$\tau$ distance for
the PFSP, the reported fitness differences between various distance metrics are
small, hence, we conjecture that the same behaviour will be observed with other
distance metrics.

@@

Table~\ref{tab:results} shows overall results over all instances evaluated (excluding the synthetic LOP instances). In particular, the mean fitness (and standard deviation) of the best solution found at the end of each run,
In terms of best fitness achieved on each run, and taking into account the
random behaviour observed in some instances and discussed above,\ekhine{}{falta un cacho aqui}\MANUEL{Mío}
